---
title: "Reproducible Research Project 1"
author: "Lacey Asplund"
date: "Friday, June 12, 2015"
output: html_document
---

##Loading & PreProcessing the data

```{r}
setwd("/Users/lacey.gray/Coursera/ReproducibleResearch/Project")
activity <-read.csv("activity.csv", header=T, sep=",")
```

##Mean Total Number of Steps Taken per Day

1. Calculate the total number of steps taken per day

```{r}
sum = tapply(activity$steps, activity$date, sum, na.rm=T)
```

2. Make a histogram of the total number of steps taken each day


```{r}
hist(sum, xlab = "Sum of Steps per Day", main="Steps per Day Historgram")
```

3. Calculate and report the mean and median of the total number of steps taken per day

The mean is :
```{r}
mean1 = round(mean(sum))
print(mean1)
```

The median is :
```{r}
median1 = round(median(sum))
print(median1)
```

##Average Daily Activity Pattern


1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
FiveMin = tapply(activity$steps, activity$interval, mean, na.rm=T)
plot(FiveMin ~ unique(activity$interval), type="l", xlab = "5 Minute Interval", main="Average Activity by Five Minute Interval")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
FiveMin[which.max(FiveMin)]
```

##Impute Missing Values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
table(is.na(activity) == TRUE)
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
<b> I chose to impute the NA's by using the mean for the 5 minute interval </b>

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activity2 <- activity
for (i in 1:nrow(activity)){
        if(is.na(activity$steps[i])){
                activity2$steps[i] <-FiveMin[[as.character(activity[i, "interval"])]]
        }
}
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

###Histogram of daily activity
```{r}
sum2 = tapply(activity2$steps, activity2$date, sum, na.rm=T)
hist(sum2, xlab = "Sum of Steps per Day", main="Steps per Day Historgram")
```

###Mean and Median steps taken per day
The mean is :
```{r}
mean2 = round(mean(sum2))
print(mean2)
```

The median is :
```{r}
median2 =round(median(sum2))
print(median2)
```


###Compare activity to activity2

```{r}
summary <- NULL
summary <- rbind(summary, data.frame(mean = c(mean1, mean2), median = c(median1, median2)))
rownames(summary) = c("with NA's", "without NA's")
print(summary)
```

To confirm no more NA's
```{r}
table(is.na(activity2) == TRUE)
```

###Impact of imputing NA's

The impact of imputing the NA's instead of removing them increases the mean and median as seen in the above.

##Differences in activity patterns between weekdays and weekends

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

###Add column containing the factor variable containing "weekday days" or "weekend days"

```{r}
activity2$weekday = c("weekday")
activity2[weekdays(as.Date(activity2[ ,2])) %in% c("Saturday", "Sunday", "samedi", "dimanche", "saturday", "sunday", "Samedi", "Dimanche"), ][4] <-c("weekend")
table(activity2$weekday == "weekend")
```

```{r}
activity2$weekday <- factor(activity2$weekday)
```

###Creating a new dataframe usable by the lattice package

```{r}
activity2_wkend <- subset(activity2, activity2$weekday == "weekend")
activity2_wkday <- subset(activity2, activity2$weekday == "weekday")

wkday_activity2_mean <-tapply(activity2_wkday$steps, activity2_wkday$interval, mean)
wkend_activity2_mean <-tapply(activity2_wkend$steps, activity2_wkend$interval, mean)
```


2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
library(lattice)
weekday <- data.frame(interval = unique(activity2_wkday$interval), avg = as.numeric(wkday_activity2_mean), day = rep("weekday", length(wkday_activity2_mean)))
weekend <- data.frame(interval = unique(activity2_wkend$interval), avg = as.numeric(wkend_activity2_mean), day = rep("weekend", length(wkend_activity2_mean)))
total <- rbind(weekday, weekend)

xyplot(avg ~ interval | day, data = total, layout = c(1, 2), 
       type = "l", ylab= "# of steps", main = "Activity patterns between Weekdays and Weekends")